MATH 245 Homework 1

### 2024-02-03 PDEs/math245_hw1.Rmd
bio hw 3

Set $\xi=xe^t$ and $\eta=x$. Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}+\pdv{u}{\xi}e^t $$

and similarly
$$u_t= \pdv{u}{t}=\pdv{u}{\eta}\pdv{\eta}{t}+\pdv{u}{\xi}\pdv{\xi}{t}=\pdv{u}{\xi}xe^t $$

Which means that 
$$u_t-xu_x=\pdv{u}{\xi}xe^t -x\left(\pdv{u}{\eta}+\pdv{u}{\xi}e^t \right)=0$$
Assuming $x\neq0$, this means
$$u_\eta=0$$
$$u=\int u_\eta \, d\eta \> + f(\xi)=0+ f(\xi)$$

So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$$
\end{tcolorbox}

@@

Now we define our new coordinate system as $\xi=y-\frac{x^2}{2}, \, \eta=x$

Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}-x\pdv{u}{\xi}$$

and similarly
$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}=$$
$$\pdv{u}{\xi}$$

$$u_x+xu_y=\pdv{u}{\eta}-x\pdv{u}{\xi}+x\pdv{u}{\xi}=\pdv{u}{\eta}$$

And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
$$\pdv{u}{\eta}=\left(y-\frac{x^2}{2}\right)^2=\xi^2$$
Now we integrate with respect to $\eta$ to find
$$u=\eta\xi^2+F(\xi)$$

At $u(0,y)$, this becomes $u=0y^2+F(y)$, or $u(0,y)=F(y)$.
On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
$$u=\eta\xi^2+e^\xi$$
Substitute back in $x$ and $y$:
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$
\end{tcolorbox}

@@

Now define our new coordinate system as $xi=\frac{x}{2}+\frac{\ln x}{2}-y, \quad \eta=x$.  
Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\xi}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\eta}$$
$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}= -\pdv{u}{\xi}$$
Plugging these into our original PDE,
$$2x\left( \pdv{u}{\xi}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\eta} \right)+\left(x+1\right)\left(-\pdv{u}{\xi}\right)=$$
$$\pdv{u}{\xi}\left(x+1\right)+2x \pdv{u}{\eta}-\pdv{u}{\xi}\left(x+1\right)=2x \pdv{u}{\eta}$$
So we must have $2x\pdv{u}{\eta}=y$. Now replace $\eta$ and $\xi$ by $x$ and $y$:
$$2\eta\pdv{u}{\eta}=\frac{\eta}{2}+\frac{\ln\eta}{2}-\xi$$

$$\pdv{u}{\eta}=\frac{1}{4}+\frac{\ln\eta}{4\eta}-\frac{\xi}{2n}$$

Now integrate with respect to $\eta$:
$$u=\frac{\eta}{4}+\frac{(\ln\eta)^2}{8}-\frac{\xi \ln\eta}{2}+F(\xi)$$
$$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)+F\left(\frac{x}{2}+\frac{\ln x}{2}-y\right)$$
Now we use the fact that $u(1,y)=2y$:
$$u(1,y)=\frac{1}{4}+F\left(\frac{1}{2}-y\right)=2y$$
$$F\left(\frac{1}{2}-y\right)=2y-\frac{1}{4}$$
So $F(\alpha)=-2\alpha+\frac{3}{4}$ for arbitrary $\alpha$. Then
$$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)-x-\ln x +2y+\frac{3}{4}$$
Which can be simplified to
$$u(x,y)=\frac{-3x}{4}-\frac{(\ln x)^2}{8}-\frac{x\ln x}{4}+\frac{y\ln x}{2}-\ln x+2y+\frac{3}{4}$$
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=-\ln x\left(\frac{x}{4} +\frac{\ln x}{8}+ \frac{y}{2} +1 \right)-\frac{3x}{4}+2y+\frac{3}{4} $$
\end{tcolorbox}

>>>>

Set $\eta=xe^t$ and $\xi=x$. Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\xi}\pdv{\xi}{x}+\pdv{u}{\eta}\pdv{\eta}{x}=$$
$$\pdv{u}{\xi}+\pdv{u}{\eta}e^t $$

and similarly
$$u_t= \pdv{u}{t}=\pdv{u}{\xi}\pdv{\xi}{t}+\pdv{u}{\eta}\pdv{\eta}{t}=\pdv{u}{\eta}xe^t $$

Which means that 
$$u_t-xu_x=\pdv{u}{\eta}xe^t -x\left(\pdv{u}{\xi}+\pdv{u}{\eta}e^t \right)=0$$
Assuming $x\neq0$, this means
$$u_\xi=0$$
$$u=\int u_\xi \, d\xi \> + f(\eta)=0+ f(\eta)$$

So we have $u(t,x)=f(\eta)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$$
\end{tcolorbox}

@@

Now we define our new coordinate system as $\eta=y-\frac{x^2}{2}, \, \xi=x$

Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\xi}\pdv{\xi}{x}+\pdv{u}{\eta}\pdv{\eta}{x}=$$
$$\pdv{u}{\xi}-x\pdv{u}{\eta}$$

and similarly
$$u_y= \pdv{u}{y}=\pdv{u}{\xi}\pdv{\xi}{y}+\pdv{u}{\eta}\pdv{\eta}{y}=$$
$$\pdv{u}{\eta}$$

$$u_x+xu_y=\pdv{u}{\xi}-x\pdv{u}{\eta}+x\pdv{u}{\eta}=\pdv{u}{\xi}$$

And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
$$\pdv{u}{\xi}=\left(y-\frac{x^2}{2}\right)^2=\eta^2$$
Now we integrate with respect to $\xi$ to find
$$u=\xi\eta^2+F(\eta)$$

At $u(0,y)$, this becomes $u=0y^2+F(y)$, or $u(0,y)=F(y)$.
On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
$$u=\xi\eta^2+e^\eta$$
Substitute back in $x$ and $y$:
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$
\end{tcolorbox}

@@

Now define our new coordinate system as $xi=\frac{x}{2}+\frac{\ln x}{2}-y, \quad \xi=x$.  
Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\xi}\pdv{\xi}{x}+\pdv{u}{\eta}\pdv{\eta}{x}=$$
$$\pdv{u}{\eta}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\xi}$$
$$u_y= \pdv{u}{y}=\pdv{u}{\xi}\pdv{\xi}{y}+\pdv{u}{\eta}\pdv{\eta}{y}= -\pdv{u}{\eta}$$
Plugging these into our original PDE,
$$2x\left( \pdv{u}{\eta}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\xi} \right)+\left(x+1\right)\left(-\pdv{u}{\eta}\right)=$$
$$\pdv{u}{\eta}\left(x+1\right)+2x \pdv{u}{\xi}-\pdv{u}{\eta}\left(x+1\right)=2x \pdv{u}{\xi}$$
So we must have $2x\pdv{u}{\xi}=y$. Now replace $\xi$ and $\eta$ by $x$ and $y$:
$$2\xi\pdv{u}{\xi}=\frac{\xi}{2}+\frac{\ln\xi}{2}-\eta$$

$$\pdv{u}{\xi}=\frac{1}{4}+\frac{\ln\xi}{4\xi}-\frac{\eta}{2n}$$

Now integrate with respect to $\xi$:
$$u=\frac{\xi}{4}+\frac{(\ln\xi)^2}{8}-\frac{\eta \ln\xi}{2}+F(\eta)$$
$$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)+F\left(\frac{x}{2}+\frac{\ln x}{2}-y\right)$$
Now we use the fact that $u(1,y)=2y$:
$$u(1,y)=\frac{1}{4}+F\left(\frac{1}{2}-y\right)=2y$$
$$F\left(\frac{1}{2}-y\right)=2y-\frac{1}{4}$$
So $F(\alpha)=-2\alpha+\frac{3}{4}$ for arbitrary $\alpha$. Then
$$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)-x-\ln x +2y+\frac{3}{4}$$
Which can be simplified to
$$u(x,y)=\frac{-3x}{4}-\frac{(\ln x)^2}{8}-\frac{x\ln x}{4}+\frac{y\ln x}{2}-\ln x+2y+\frac{3}{4}$$
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=-\ln x\left(\frac{x}{4} +\frac{\ln x}{8}+ \frac{y}{2} +1 \right)-\frac{3x}{4}+2y+\frac{3}{4} $$
\end{tcolorbox}
